---
title: "The relationship between location of breach and the number of individuals affected"
author: "Tanu Roy"
date: "5/3/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```


```{r}
breach_data <- read_csv('C:\Users\tanu roy\OneDrive\Desktop\SYS2202\cyber-security-final',
                     col_types = cols(
                      Location_of_Breach = col_factor(),
                      Individuals_Affected = col_integer()
                     )
                  )
```
```{r}
head(breach_data)
```

### 3.1.3 Location of Breach Variable 


#### 3.1.2.1 Visualising distributions (Barcharts, Histograms) (5 points)

```{r}

```


**- Which values are the most common? Why?** 


**- Which values are rare? Why? Does that match your expectations?** 


**- Can you see any unusual patterns? What might explain them?**


**- Are there clusters in the data? If so,**


  **- How are the observations within each cluster similar to or different from each other?**

  
  **- How can you explain or describe the clusters?**


  
#### 3.1.2.2 Unusual values (2 points) 

**- Describe and demonstrate how you determine if there are unusual values in the data. E.g. too large, too small, negative, etc.**

```{r}

```


**- Describe and demonstrate how you determine if they are outliers.**



```{r}

```

**- Show how do your distributions look like with and without the unusual values.**


```{r}

```

**- Discuss whether or not you need to remove unusual values and why.**


#### 3.1.2.3 Missing values (2 points)

**- Does this variable include missing values? Demonstrate how you determine that.**



```{r}

```

**- Demonstrate and discuss how you handle the missing values. E.g., removing, replacing with a constant value, or a value based on the distribution, etc.**



**- Show how your data looks in each case after handling missing values.Describe and discuss the distribution.** 


 

#### 3.1.2.4 Does converting the type of this variable help exploring the distribution of its values or identifying outliers or missing values? (3)
 

**- What type can the variable be converted to?**



```{r}

```


**- How will the distribution look? Please demonstrate with appropriate plots.**



```{r}

```

#### 3.1.2.5 What new variables do you need to create?  (3)


**- List the variables**


```{r}

```

```{r}

```


```{r}

```

```{r}

```


**- Describe and discuss why they are needed and how you plan to use them.**




```{r}

```
### 3.1.3 Individuals_affected Variable  


#### 3.1.2.1 Visualising distributions (Barcharts, Histograms) (5 points)

```{r}

```


**- Which values are the most common? Why?** 



**- Which values are rare? Why? Does that match your expectations?** 



```{r}

```


**- Can you see any unusual patterns? What might explain them?**


**- Are there clusters in the data? If so,**


  **- How are the observations within each cluster similar to or different from each other?**
  
 
  
  **- How can you explain or describe the clusters?**

 
  
#### 3.1.2.2 Unusual values (2 points) 

**- Describe and demonstrate how you determine if there are unusual values in the data. E.g. too large, too small, negative, etc.**


```{r}

```


**- Describe and demonstrate how you determine if they are outliers.**


```{r}

```

**- Show how do your distributions look like with and without the unusual values.**


```{r}

```

**- Discuss whether or not you need to remove unusual values and why.**

#### 3.1.2.3 Missing values (2 points)

**- Does this variable include missing values? Demonstrate how you determine that.**



```{r}

```

**- Demonstrate and discuss how you handle the missing values. E.g., removing, replacing with a constant value, or a value based on the distribution, etc.**

**- Show how your data looks in each case after handling missing values.Describe and discuss the distribution.** 

 

#### 3.1.2.4 Does converting the type of this variable help exploring the distribution of its values or identifying outliers or missing values? (3)



**- What type can the variable be converted to?**


```{r}

```


**- How will the distribution look? Please demonstrate with appropriate plots.**


```{r}

```

#### 3.1.2.5 What new variables do you need to create?  (3)


**- List the variables**


**- Describe and discuss why they are needed and how you plan to use them.**



## 3.2. What type of covariation occurs between the variables? (30 points)
If you don't have variables of a certain type in the original dataset or among the created variables (features), you can further create them from the existing variables. See RDS chap. 5, 7.5 and 7.6.

### 3.2.1 Between a categorical and continuous variable (10 points)

**- Describe what type of visualization you can use and why.**

```{r fig.height=10, fig.align= "center"}

```


**- Describe the patterns and relationships you observe. Could the identified patterns be due to coincidence (i.e. random chance)?**


**- Describe the relationship implied by the pattern? (e.g., positive or negative correlation)**


**- Calculate the strength of the relationship implied by the pattern (e.g., correlation)**



```{r}

```


**- Discuss what other variables might affect the relationship**
  


**- Does the relationship change if you look at individual subgroups of the data? Please discuss and demonstrate.**



```{r fig.height=10, fig.align= "center"}

```


**- Demonstrate if converting the type of these variables help exploring the relationship.**


```{r}

```


```{r}

 
```


```{r}

```

```{r}

```


```{r}

```


```{r}

```


**- Discuss how the observed patterns support/reject your hypotheses or answer your questions.**
